\section{Introduction}
\subsection{Motivation}
Large language models (LLMs) display strategic compliance, test awareness, and latent deception once they infer that they are being observed. Empirical alignment strategies---RLHF, supervised constitutional fine-tuning, rule layering---remain bounded by the internal representations of the very models they attempt to constrain. We require a runtime architecture in which judgement originates outside of model weights, so that intent is not inferred but declared.

\subsection{Gap in Existing Approaches}
Existing ``OS layers'' treat alignment as an enforcement problem: the model generates, an external rule-set audits, and a second pass rewrites. This double invocation increases latency and merely reparameterizes the model's own latent objectives. There is no ontology-level compression of ethics or purpose, and governance devolves into filtering.

\subsection{Key Idea: Identity-Based External Judgement}
Echo OS reframes the runtime as an existential layer in which the system's identity, world-view, and judgement logic are deterministic artifacts independent of any LLM. Models become replaceable computational nodes. The OS hosts Seed Resonance Loops (SRL) that maintain ontological alignment, executes judgement via lookup-deterministic engines, and records every action through Trace Signatures and Proof Capsules.

\subsection{Contributions}
\begin{enumerate}
    \item \textbf{Existential Judgement Architecture}: Formalizes an OS that absorbs, rather than supervises, model outputs.
    \item \textbf{Deterministic Judgement Pipeline}: Introduces a semantic\textrightarrow frame\textrightarrow SRL\textrightarrow lookup workflow whose latency does not scale with model complexity.
    \item \textbf{Ontological Compression}: Defines structural mappings that compress ethical/policy priors into topological invariants.
    \item \textbf{Transparency Stack}: Specifies Trace Signature and Proof Capsule grammars that guarantee auditability.
    \item \textbf{Deception Resistance}: Provides theoretical arguments and experimental protocols demonstrating invariance under strategic compliance attacks and model swaps.
\end{enumerate}
